---
title: "Genotype Simulation - A Mini-How-To"
author: "Peter von Rohr"
date: "`r Sys.Date()`"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Disclaimer
This mini-how-to describes possible approaches in order to simulate genotypes in a pedigree. 

## Other packages
Knowing that Karl Broman has authored a QTL mapping package and a book on QTL-mapping, a quick search for other packages resulted in the package `QTLRel`. 


```{r InstallQTLRel}
if (!require(QTLRel))
  install.packages(pkgs = "QTLRel")
```

The package comes with an example dataset that contains a pedigree, genetic data and phenotypic information. The data can be made available using

```{r LoadQLTRelData}
load(file = file.path(system.file(package = "QTLRel", "data"), "miscEx.RData"))
```

### Estimation of a-Effects using phenotypic data
In existing data, it is difficult to find the six most significant loci. Alternatively, we start six loci and simulate a new phenotype.

```{r SelectLoci}
phenDatAEst <- data.frame(sex = pdatF8[,"sex"], age = pdatF8[,"age"], gdatF8[,5:10])
str(phenDatAEst)
```

Simulating the genetic part for a single individual corresponds to 

```{r SimSingleIndividual}
 a_values <- c(13,3,5.3,2.1,6.7,9.1)
crossprod(as.numeric(phenDatAEst[1,3:8]-2),a_values)
```



## Goal
Given a pedigree, a number of loci and their corresponding allele frequencies in the base population, we want to distribute the genotypes over the pedigree such that mendelian rules of inheritance are satisfied.


## Implementation
The pedigree simulation is implemented using two basic classes of objects

1. A class of GenoLocus objects which describe a genetic locus on the population level. This object contains the name of the locus, the different alleles and the allele frequencies. 
2. A class of a gene pool object which contains a collection of GenoLocus objects. 
3. A class of individuals that describe the genetic composition of single individuals.


